VM-Centric Snapshot Deduplication with Fault Isolation for Converged Cloud Architectures

### 2014-01-31 snapshot_system/paper_analysis.tex

\begin{figure}
    \centering
    \subfigure[Sharing of file system blocks under VC]
    {
        \includegraphics[width=3in]{images/share_vc}
        \label{fig:share_vc}
    }
    \\
    \subfigure[Sharing of file system blocks under VO]
    {
        \includegraphics[width=3in]{images/share_vo}
        \label{fig:share_vo}
    }
    \caption{Bipartite association of VMs and file system blocks under (a) VC and (b) VO. }
    \label{fig:share}
\end{figure}

@@

As illustrated in Figure~\ref{fig:share}, we build a bipartite graph representing the
association from unique file system blocks to their corresponding VMs in two 
approaches. 
%We use a file system block rather than data chunk as our unit of failure
%because the underlying file system keeps file system blocks as its base unit of storage, and in
%the case of 64MB file system blocks  and 4KB chunks, there are 16K chunks per file block. For VM centric,
%we assign extra replication to the PDS blocks, while for VM oblivious, the replication
%degree is fixed.
For VC, each VM has an average number $N_1$ of non-PDS file system blocks
and  has  an average of  $N_2$ PDS file system blocks. 
Each non-PDS block is associated with only one VM.
% and  we denote that PDS blocks are shared by an average of $V_c$ VMs. 
Then by counting outgoing edges from VMs in Figure~\ref{fig:share}(a),
\[
n* N_1   = \mbox{ Number of non-PDS file system blocks in VC.} 
\]

@@

Since we choose 2-4\% of unique chunks for PDS
and Section~\ref{sect:evaldedup} 
shows that the deduplication efficiency of VC is very close to that of VO,
the number of non-PDS file blocks in VC is approximately close to the number of file blocks in VO.
Then
\[
\frac{N_o}{N_1} \approx  V_o.
\]
%Thus $N_o >> N_1$.

@@

\noindent $r \le d<r_c$:  In this case there is no PDS data loss and we need to look at the non-PDS data loss. 
The full snapshot availability of a VM is: 
%\begin{equation}
%\label{eq:VC1}
%(1-\frac{\binom{d}{r}} { \binom{p}{r} })^{N_1}.
\[
A(r)^{N_1}.
\]
%\end{equation}
Since $N_1$ is typically much smaller than $N_o$, 
the VC approach has a higher availability of VM snapshots than VO in this case.
%In the evaluation discussed in Section~\ref{sect:evaldedup}, 
%We have considered a worst case scenario where
%every PDS FSB is shared by all VMs in the VC approach, which leads a large $N_2$ value. 
%Even with $N_2$ being much higher than $N_o$ (as discussed below),
%the availability of VC snapshots is much higher than VO.

\noindent $r_c \leq d$: Both non-PDS and PDS file system blocks in VC can have a loss.
The full snapshot availability of  a VM in the VC approach is
%\begin{equation}
%\label{eq:VC2}
% (1-\frac{ \binom{d}{r}} { \binom{p}{r} })^{N_1} 
% *
% (1-\frac{ \binom{d}{r_c}} { \binom{p}{r_c} })^{N_2}.
\[
A(r)^{N_1} * A(r_c)^{N_2}.
\]
That is still smaller than that of $V_O$ based on the observation with our data. There are two reasons for this:  
1) $N_1$ is much smaller than $N_o$ and we are observing that $N_1+N_2<N_o$. 
2)  $A(r) < A(r_c)$ because $r < r_c$.  
Table~\ref{tab:fsb-availability} lists the $A(r)$ values with
%that the availability of an individual file system block
different replication degrees, to demonstrate the gap between  $A(r)$ and  $A(r_c)$.
%\end{equation}
%comparing  Formula~\ref{eq:VO} and~\ref{eq:VC}. 
%In the evaluation discussed in Section~\ref{sect:evaldedup},
%We have considered a worst case scenario that
%every PDS FSB is shared by all VMs in the VC approach, which leads to a large $N_2$ value. 
%Even with that, the availability of VC snapshots is still much higher than VO and

>>>>

\begin{figure}[htb]
    \centering
    \subfigure[Sharing of file system blocks under VC]
    {
        \includegraphics[width=3in]{images/share_vc}
        \label{fig:share_vc}
    }
    \\
    \subfigure[Sharing of file system blocks under VO]
    {
        \includegraphics[width=3in]{images/share_vo}
        \label{fig:share_vo}
    }
    \caption{Bipartite association of VMs and file system blocks under (a) VC and (b) VO. }
    \label{fig:share}
\end{figure}

@@

As illustrated in Figure~\ref{fig:share}, we build a bipartite graph representing the
association from unique file system blocks to their corresponding VMs in two 
approaches. 
%We use a file system block rather than data chunk as our unit of failure
%because the underlying file system keeps file system blocks as its base unit of storage, and in
%the case of 64MB file system blocks  and 4KB chunks, there are 16K chunks per file block. For VM centric,
%we assign extra replication to the PDS blocks, while for VM oblivious, the replication
%degree is fixed.
For VC, each VM has an average number $N_1$ of non-PDS file system blocks
and  has  an average of  $N_2$ PDS file system blocks. 
Each non-PDS block is associated with only one VM.
% and  we denote that PDS blocks are shared by an average of $V_c$ VMs. 
Then by counting outgoing edges from VMs in Figure~\ref{fig:share}(a), we get:
\[
n* N_1   = \mbox{ Number of non-PDS file system blocks in VC.} 
\]

@@

Since we choose 2-4\% of unique chunks for PDS
and Section~\ref{sect:evaldedup} 
shows that the deduplication efficiency of VC is very close to that of VO,
the number of non-PDS file blocks in VC is similar to the number of file blocks in VO.
Then
\[
\frac{N_o}{N_1} \approx  V_o.
\]
%Thus $N_o >> N_1$.

@@

\noindent {\bf $r \le d<r_c$:}  In this case there is no PDS data loss and we need to look at the non-PDS data loss. 
The full snapshot availability of a VM is: 
%\begin{equation}
%\label{eq:VC1}
%(1-\frac{\binom{d}{r}} { \binom{p}{r} })^{N_1}.
\[
A(r)^{N_1}.
\]
%\end{equation}
Since $N_1$ is typically much smaller than $N_o$, 
the VC approach has a higher availability of VM snapshots than VO in this case.
%In the evaluation discussed in Section~\ref{sect:evaldedup}, 
%We have considered a worst case scenario where
%every PDS FSB is shared by all VMs in the VC approach, which leads a large $N_2$ value. 
%Even with $N_2$ being much higher than $N_o$ (as discussed below),
%the availability of VC snapshots is much higher than VO.

\noindent {\bf $r_c \leq d$:} Both non-PDS and PDS file system blocks in VC can have a loss.
The full snapshot availability of  a VM in the VC approach is
%\begin{equation}
%\label{eq:VC2}
% (1-\frac{ \binom{d}{r}} { \binom{p}{r} })^{N_1} 
% *
% (1-\frac{ \binom{d}{r_c}} { \binom{p}{r_c} })^{N_2}.
\[
A(r)^{N_1} * A(r_c)^{N_2}.
\]
That is still smaller than that of $V_O$ based on the observations of our data. There are two reasons for this:  
1) $N_1$ is much smaller than $N_o$ and we are observing that $N_1+N_2<N_o$. 
2)  $A(r) < A(r_c)$ because $r < r_c$.  
Table~\ref{tab:fsb-availability} lists the $A(r)$ values with
%that the availability of an individual file system block
different replication degrees, to demonstrate the gap between  $A(r)$ and  $A(r_c)$.
%\end{equation}
%comparing  Formula~\ref{eq:VO} and~\ref{eq:VC}. 
%In the evaluation discussed in Section~\ref{sect:evaldedup},
%We have considered a worst case scenario that
%every PDS FSB is shared by all VMs in the VC approach, which leads to a large $N_2$ value. 
%Even with that, the availability of VC snapshots is still much higher than VO and

### 2014-01-31 snapshot_system/paper_background.tex

Figure~\ref{fig:collocated} illustrates a converged IaaS cloud architecture 
where
each commodity server hosts a number of virtual machines and storage of these servers
is clustered using a distributed file system~\cite{googlefs03,hdfs10,NutanixPaper}.
Each physical machine hosts multiple virtual machines.  Every virtual machine
runs its own guest operating system (typically one of only a few
installed on the system) and accesses virtual hard disks 
stored as image files maintained by the operating system running on the
physical host.
For VM snapshot backup, file-level semantics are normally not provided.
Snapshot operations take place at the virtual device driver level, which
means no fine-grained file system metadata can be used to determine the changed data. 
%This architecture does not use a separate  backup facility for a lower cost and 
%reduces the network bandwidth consumption in transferring the raw data for backup. 
%A backup service may co-locate with other cloud services on these commodity servers and 
%stores deduplicated data  in  the storage in this cluster or in
%a separate storage cluster.

@@

File backup systems have been developed to use fingerprints generated for
data ``chunks''  to identify duplicate
content~\cite{venti02,Rhea2008}.  In its simple implementation,
it is expensive to compare a large number of 
fingerprints
so several techniques have been proposed to improve duplicate identification. 
For example, the data domain method ~\cite{bottleneck08} 
uses  an in-memory Bloom filter and a prefetching cache for data chunks 
which may be
accessed.  
%An improvement to this work with parallelization is in ~\cite{MAD210,DEBAR}.
Approximation techniques are studied in~\cite{extreme_binning09,Guo2011,WeiZhangIEEE}  
to reduce memory requirements at the cost of some loss of deduplication
efficiency.
Additional inline deduplication techniques are studied in ~\cite{sparseindex09,Guo2011,Srinivasan2012}
and a parallel batch solution for cluster-based deduplication is 
studied in ~\cite{wei2013}. 
All of the above approaches have focused on optimization of deduplication
efficiency, and none of them have considered the impact
of deduplication on fault resilience in a cloud environment.  To do so, we
examine the following properties of VM snapshot deduplication for converged
IaaS architectures.
%considered in this paper.
%Our consideration is discussed below.
%VM dependence minimization during deduplication and file system block management.
\paragraph*{Local versus Global Deduplication} --
If deduplication is implemented strictly in the storage system,
a data chunk is compared with fingerprints collected from all VMs during
the deduplication process, and only a small number of
copies of duplicates are stored in the storage.  This sharing
artificially creates hidden data dependencies among different VMs owned by
different 
who
contract for isolation 
by the IaaS cloud platform.  That is, the unit of failure visible to the cloud
user is the VM, but the unit of sharing in a deduplicated data store is the
data chunk.

@@

Indeed, the greater the deduplication efficiency, the less
the data redundancy that is maintained, and the greater the ``hidden'' data
dependence between VMs that must be hosted as if they were physically isolated.
%machine failures happen periodically 
%in a large-scale cloud and
%loss of a small number of shared data chunks can 
%cause the unavailability of snapshots for a large number of virtual machines.
Thus,
restricting deduplication to be local to each VM (but across snapshots) 
aligns the units of storage with the isolation properties guaranteed by the
cloud at the possible expense of storage efficiency.
%Localizing the impact of deduplication can increase fault isolation and resilience.
%Thus from the fault tolerance point of view,  duplicate sharing among multiple VMs is 
%discouraged.

Additionally, storage efficient deduplication can be computationally expensive
to implement in a distributed setting.  In particular, duplicate detection is
logically a ``global'' operation with respect to the population of data chunks
under consideration.
By localizing duplicate search, the computational requirements
for deduplication can be reduced, because fewer chunks must be compared.
Further, deduplication of otherwise unrelated VMs can be implemented in
parallel.

Another disadvantage of cross-VM sharing is that it complicates 
snapshot deletion, 
which occurs frequently when snapshots expire regularly. 
The mark-and-sweep approach~\cite{Guo2011,Fabiano2013}  is effective for
deletion, but still carries a significant cost, again in the distributed cloud
setting.
%to count if a data chunk is still shared by other snapshots.

@@

%minimize data sharing and simplify deletion while sacrificing 
%deduplication efficiency, and  can facilitate parallel execution of snapshot operations.
\paragraph*{Units of Storage versus Units of Sharing} --
The file system block  size in a distributed file system such as  
the Hadoop file system and the Google file system~\cite{googlefs03}
is large (e.g.  64MB) so as to optimize file system performance
for big data.
Typically, deduplication implementations use smaller block sizes (4K bytes to
8K bytes on average~\cite{Jin2009}).
In addition to the failure correlation that cross-VM data deduplication can
introduce, disk block sharing of data chunks is another potential source of
hidden data dependence.
Thus a VM-centric solution should consider how data chunks from VMs are
distributed among large file system blocks when a large-scale distributed file
system is used for disk image files.

@@

For the remainder of this paper, 
we term the traditional deduplication approach {\em VM-oblivious} (VO)
because it manages duplicate data chunks without consideration of VMs.
With the above  considerations in mind, we propose a 
VM-centric approach (called VC)
for a co-located backup service that has a resource usage profile
suitable for use with converged cloud architecture.

@@

We will first discuss and analyze the integration of the VM-centric deduplication strategies with fault isolation, 
and discuss snapshot deletion support, then present an implementation design.

>>>>

Figure~\ref{fig:collocated} illustrates a converged IaaS cloud architecture 
where
each commodity server hosts a number of virtual machines and storage of these servers
is clustered using a distributed file system~\cite{googlefs03,hdfs10,NutanixPaper}.
Each physical machine hosts multiple virtual machines.  Every virtual machine
runs its own guest operating system and accesses virtual hard disks 
stored as image files maintained by the operating system running on the
physical host.
For VM snapshot backup, file-level semantics are normally not provided.
Snapshot operations take place at the virtual device driver level, which
means no fine-grained file system metadata can be used to determine the changed data. 
%This architecture does not use a separate  backup facility for a lower cost and 
%reduces the network bandwidth consumption in transferring the raw data for backup. 
%A backup service may co-locate with other cloud services on these commodity servers and 
%stores deduplicated data  in  the storage in this cluster or in
%a separate storage cluster.

@@

File backup systems have been developed to use fingerprints generated for
data ``chunks''  to identify duplicate
content~\cite{venti02,Rhea2008}.  In a simple implementation,
it is expensive to compare a large number of 
fingerprints
so several techniques have been proposed to improve duplicate identification. 
For example, the data domain method ~\cite{bottleneck08} 
uses  an in-memory Bloom filter and a prefetching cache for data chunks 
which may be
accessed.  
%An improvement to this work with parallelization is in ~\cite{MAD210,DEBAR}.
Approximation techniques are studied in~\cite{extreme_binning09,Guo2011,WeiZhangIEEE}  
to reduce memory requirements at the cost of some loss of deduplication
efficiency.
Additional inline deduplication techniques are studied in ~\cite{sparseindex09,Guo2011,Srinivasan2012}
and a parallel batch solution for cluster-based deduplication is 
studied in ~\cite{wei2013}. 
All of the above approaches have focused on optimization of deduplication
efficiency or speed, and none of them have considered the impact
of deduplication on fault resilience in a cloud environment.  To do so, we
examine the following properties of VM snapshot deduplication for converged
IaaS architectures.
%considered in this paper.
%Our consideration is discussed below.
%VM dependence minimization during deduplication and file system block management.
\paragraph*{Local versus Global Deduplication} --
If deduplication is implemented strictly in the storage system,
a data chunk is compared with fingerprints collected from all VMs during
the deduplication process, and only a small number of
copies of duplicates are stored in the storage.  This sharing
artificially creates hidden data dependencies among different VMs owned by
different 
who
contract for isolation 
by the IaaS cloud platform.  That is, the unit of failure visible to the cloud
user is the VM, but the unit of sharing in a deduplicated data store is the
data chunk.

@@

Indeed, the greater the deduplication efficiency, the less
the data redundancy is maintained, and the greater the ``hidden'' data
dependence between VMs that must be hosted as if they were physically isolated.
%machine failures happen periodically 
%in a large-scale cloud and
%loss of a small number of shared data chunks can 
%cause the unavailability of snapshots for a large number of virtual machines.
Thus,
restricting deduplication to be local to each VM (but across snapshots) 
aligns the units of storage with the isolation properties guaranteed by the
cloud at the possible expense of storage efficiency.
%Localizing the impact of deduplication can increase fault isolation and resilience.
%Thus from the fault tolerance point of view,  duplicate sharing among multiple VMs is 
%discouraged.

Additionally, storage efficient deduplication can be computationally expensive
to implement in a distributed setting.  In particular, duplicate detection is
logically a ``global'' operation over all data chunks
under consideration.
By localizing duplicate search, the computational requirements
for deduplication can be reduced, because fewer chunks must be compared.
Further, deduplication of VMs can easily be implemented in
parallel.

Another disadvantage of cross-VM sharing is that it complicates 
snapshot deletion, 
which occurs frequently when snapshots expire regularly. 
The mark-and-sweep approach~\cite{Guo2011,Fabiano2013}  is effective for
deletion, but still carries a significant cost, especially in the distributed cloud
setting.
%to count if a data chunk is still shared by other snapshots.

@@

%minimize data sharing and simplify deletion while sacrificing 
%deduplication efficiency, and  can facilitate parallel execution of snapshot operations.
\paragraph*{Units of Storage versus Units of Sharing} --
The file system block  size in a distributed file system such as  
the Hadoop file system and the Google file system~\cite{googlefs03}
is large (e.g.  64MB) so as to optimize file system performance
for big data.
Typically, deduplication implementations use smaller block sizes (4K bytes to
8K bytes on average~\cite{Jin2009}).
In addition to the failure correlation that cross-VM data deduplication can
introduce, disk block sharing of data chunks is another potential source of
hidden data dependence.
Thus a VM-centric solution should consider how data chunks from VMs are
distributed among large file system blocks when a large-scale distributed file
system is used for snapshot storage.

@@

For the remainder of this paper, 
we term the traditional deduplication approach {\em VM-oblivious} (VO)
because it manages duplicate data chunks without consideration of VMs.
With the above  considerations in mind, we propose a 
VM-centric approach (called VC)
for a co-located backup service that has a resource usage profile
suitable for use with converged cloud architectures.

@@

We will first discuss and analyze the integration of the VM-centric deduplication strategies with fault isolation 
and discuss snapshot deletion support, then present an implementation design.

### 2014-01-31 snapshot_system/paper_dedup.tex

Since the best deduplication uses a non-uniform chunk size 
in the average of 4KB or 8KB~\cite{Jin2009},
we conduct additional local similarity guided deduplication on a snapshot by comparing
chunk fingerprints of a dirty segment 
with those in  {\em similar} segments from its parent snapshot. 
We define two segments are similar if their content signature is the same.
This segment content signature value is defined as the minimum value of all its chunk fingerprints 
computed during backup and is recorded in the snapshot metadata (called recipe). Note that this definition of
content similarity is an approximation~\cite{resemblance97}.  When processing a dirty segment,
its  similar segments can be found easily from the
parent snapshot recipe.  Then recipes of the similar segments are loaded to memory,
which contain chunk fingerprints to be compared.
To control the time cost of search, we set a limit on the number of  similar segments recipes to be fetched. 
% Then,
%given a set of data chunks within a dirty segment,  we compare  these chunk fingerprints
%with those in similar segments.  
For example, assume that  a segment is of size  2MB, 
its segment recipe is roughly 19KB which contains about 500 chunk fingerprints and other chunk metadata,
by limiting at most 10 similar segments to search, the amount of memory for maintaining those 
similar segment recipes is small.
As part of our local duplicate search we also compare the current segment
against the parent segment at the same offset.

@@

Since $k$ is relatively small and these top $k$ chunks are shared among multiple VMs, 
we can afford to provide extra replicas for these popular chunk data to enhance the fault resilience.
We have studied the distribution of chunk popularity~\cite{WeiZhangIEEE} 
based on a data trace from Alibaba with over 2500 VMs.
The distribution is zipf-like and popular chunks are dominating the large portion of data chunks.
We denote $\sigma$ as the percentage of unique data chunks belonging to PDS and from the evaluation we find that
$\sigma$ with a range of 1 to 4\% can deliver a fairly competitive deduplication efficiency.

>>>>

Since the best deduplication uses a non-uniform chunk size 
with an average of 4KB or 8KB~\cite{Jin2009},
we conduct additional local similarity guided deduplication on a snapshot by comparing
chunk fingerprints of a dirty segment 
with those in  {\em similar} segments from its parent snapshot. 
We define two segments are similar if their content signature is the same.
This segment content signature value is defined as the minimum value of all its chunk fingerprints 
computed during backup and is recorded in the snapshot metadata (called recipe). Note that this definition of
content similarity is an approximation~\cite{resemblance97}.  When processing a dirty segment,
its  similar segments can be found easily from the
parent snapshot recipe.  Then recipes of the similar segments are loaded to memory,
which contain chunk fingerprints to be compared.
To control the time cost of search, we set a limit on the number of  similar segment recipes to be fetched. 
% Then,
%given a set of data chunks within a dirty segment,  we compare  these chunk fingerprints
%with those in similar segments.  
For example, assume that  a segment is of size  2MB, 
its segment recipe is roughly 19KB which contains about 500 chunk fingerprints and other chunk metadata.
By limiting at most 10 similar segments to search, the amount of memory for maintaining those 
similar segment recipes is small.
As part of our local duplicate search we also compare the current segment
against the parent segment at the same offset.

@@

Since $k$ is relatively small and these top $k$ chunks are shared among multiple VMs, 
we can afford to provide extra replicas for these popular chunk data to enhance the fault resilience.
We have studied the distribution of chunk popularity~\cite{WeiZhangIEEE} 
based on a data trace from Alibaba with over 2500 VMs.
The distribution is zipf-like and popular chunks dominate the distribution.
We denote $\sigma$ as the percentage of unique data chunks belonging to PDS and from the evaluation we find that
$\sigma$ with a range of 1 to 4\% can deliver a fairly competitive deduplication efficiency.
